\begin{agradecimentos}[Acknowledgements]
	I would like to acknowledge the following people for their support in the
	development of this thesis.

	\imprimirorientador{} by his confidence and trust in myself and in this work, and by the
	opportunity he gave me to pursuit these ideas.
	\imprimircoorientador\ for his kind support at a time where this work was in its
	infancy.
	I thank both for their support, kindness, advice and mentoring.

	The thesis committee for taking the time and effort to read, evaluate and
	discuss the present thesis.

	To all professors in the Chemistry Department and in the Graduate Programme, and to the Federal University of
	Santa Catarina as a whole
	for the high-quality education and excellent work.
	A special thanks goes to Prof.\ Dr.\ Josiel D. Barbosa and
	Dr.\ Bernardo de Souza
	for their enthusiasm and outstanding collaboration, advice and support.

	To all the members of our research group, colleagues and friends at GEEM for
	their support.
	I would like to thank particularly the following people:
	Alechania Misturini,
	Alexandre Osmar Ortolan,
	Denner Ferreira,
	Letícia Maria Pequeno,
	Marina Briese,
	Matheus Colaço,
	Rômulo Peixoto,
	Vinícius Glitz,
	and,
	Vinícius Port.
	Without you I could not have made until here.

	To Scheila and Manssur for helping me when I most needed it.

	To my parents (Edson Schneider and Laureci Silveira de Souza Schneider)
	for their love and support in all these years.

	To Deise Munaro for being so kind and special,
	and for standing on my side for so long and in such difficult times.

	I am also grateful for the funding and resources provided by the following institutions:
	CNPq, for a~PhD scholarship (grant number 140485/2017--1);
	and FAPESC, FAPESP, and CESUP/UFRGS, whose valuable resources make, directly or
	indirectly, our computing facilities
	viable.

	To the developers of software I have used throughout these years, in special for
	the developers of the following projects:
	Julia,
	NumPy/SciPy,
	Vim/Neovim,
	\LaTeX\ and abn\TeX{},
	and,
	cclib.

\end{agradecimentos}
